```{r echo=F, warning=FALSE, error=FALSE}
opts_chunk$set(echo=TRUE, warning=FALSE, message=F, error=FALSE, textwidth=200)
```

A Brief Introduction To R
========================================================

In the first hour we'll do a brief tour of the R programming language. 

R is the most popular programming language for statistics and data analysis.

R is an open-source implementation of the programming language S. S was developed by John Chambers at Bell Laboratories in 1976.

The thing that makes R so powerful is the enormous number of libraries that exist for it. We'll use some of them later.

Using R
------------------
R is most often used interactively. That means that we care more about the code written during an R session is not saved, but only the output (e.g. graphics). (Although, to create reproducible research it might make sense to keep the code.)

There are two primary ways of using R:
- The command line
- RStudio (open-source IDE)

RStudio is generally more pleasant and powerful.

First Steps
------------------

```{r}
5
print(10)
6 + 7
```

An important thing is to be in the right working directory. 

In RStudio, you can set the working directory under 'Session > Set Working Directory'

Or you can set it directly in the console: 

```{r}
getwd()
setwd('/Users/briancrain/Documents/quantified_self/R_WORKSHOP/measured_me_rws/')
```

If you don't know something, the help function is very useful.
```{r}
help(mean)
```

There is also a shortcut.
```{r}
?mean
```

Seeing examples is useful too.
```{r}
example(mean)
```


Functions
------------------
Functions are extremely important in R.

They are simple to call.

```{r}
print('Hello')
```

Let's generate a random number.
```{r}
rnorm(1)
```

Or two.
```{r}
rnorm(2)
```

We can also take the mean of two random numbers.

```{r}
mean(rnorm(2))
```

Or we can save the random numbers first and then take the mean.

```{r}
v <- rnorm(2)
mean(v)
```

There two main ways to save variables in R.

```{r}
v <- rnorm(2)
```
This is also possible: rnorm(2) -> v

Calling variables is simple.
```{r}
v
```

Something more fun.

```{r}
a <- rnorm(10)
hist(a)
```

You could also do it like this.
```{r}
hist(rnorm(10))
```

Before we move on, let's write our own function. 
```{r}
f <- function(a){
  a + 1
}
f(5)
```
R functions always return the last line, but you could also use return explicitly.
```{r}
f <- function(a){
  return(a+1)
  print('I never get called')
}
f(5)
```

Data Types
------------------
### Vectors

The most fundamental data type in R is the vector.

This is the standard way of creating a vector:
```{r}
a <- c(7) 
```
But this is really a vector too:
```{r}
7 
typeof(7) == typeof(c(7))
```

You can have lots of elements in a vector. 
```{r}
c(7, 4, 9)
c(1:100)
```

But they need to be of the same type
```{r}
c(7, 'dog')
```
As you can see R coeerces both into a string here so the type is the same.

Here is how you access elements of a vector.
```{r}
a <- 1:100
a[1]
a[56]
a[3*1:25]
a[a%%3 == 0]
```
Note: Indexing starts at 1 in R, not at 0 like in python, javascript, etc.

What happens when you multiply a vector?
```{r}
a*2
```
Operations are always applied element-wise. This is very important.

### Lists

We'll just cover lists very briefly. They can contain different data types.
```{r}
L <- list(name='Sam', age=7)
```
Elements can be accessed by position or name.
```{r}
L$name
L[[2]]
```

### Data Frames

Data frames are the data type you will work with the most. The graphics package ggplot2, for example, only accepts data frames.

A data frame is a list with class 'data.frame'. Works a bit like a table in a database or like an excel sheet.

Let's create one.
```{r}
data.frame(home='Florida', income=10000)
(accountant <- data.frame(home=c('Florida', 'Berlin'), income=c(10000, 23456), stringsAsFactors = FALSE))
```

Let's add a row.
```{r}
accountant[3,] <- c('Hungary', 43)
accountant
```

Let's look at the mean income.
```{r}
mean(accountant$income)
```




You can also embed plots, for example:

```{r fig.width=7, fig.height=6}
plot(cars)
```

