```{r echo=F, warning=FALSE, error=FALSE}
opts_chunk$set(echo=TRUE, warning=FALSE, message=F, error=FALSE, textwidth=200)
```

Exploration of the Measured Me dataset
======================================

Data provided by [Measured Me](http://measuredme.com/). Thanks!

This worksheet was created in R Markdown. Markdown is a simple formatting syntax for authoring web pages. You can load up this worksheet in R Studio to run the examples directly and explore yourself!

## Metadata

A description of the dataset from the author, Konstantin:

1. Sleep duration and efficiency were tracked using Sleep Time app
2. Calorie expenditure, moderate and vigorous activity (in minutes), and steps were tracked using Bodymedia
3. Physical energy, stress, emotions positivity and intensity, happiness and life satisfaction were tracked using 10-point sliding scale in rTracker app.  
4. Physical Health was measured using 5-point scale (no symptoms= 5, ignorable symptoms = 4, distracting symptoms = 3, debilitating symptoms = 2, sickness = 1). 
5. The mental alertness was tracked using Mind Metrics app; I then converted the reaction time to Alertness score using formula 10000/alertness_time. 
6. The cognitive performance was tracked using Stroop Effect app. I converted time of responses to Cognitive score using formula 100/response_time.

## Load the data

Read in the csv file.
```{r}
mm <- read.csv("../measured_me.csv")
```

Let's take a look at the first few rows.
```{r}
head(mm)
```

It appears that the first column is a date, in an odd format.
```{r}
str(mm)
```

Yes, R isn't recognising it as a date. R thinks it's a categorical variable, or `factor`.

We can parse it so that R may convert it into the `Date` data type. There's a nice library to make this a bit easier. It also looks like the year is missing so we'll add that too.
```{r}
library(lubridate)
mm$Date <- parse_date_time(mm$Date,"md")
year(mm$Date) <- 2013
```

Then we can check to see if it worked.
```{r}
class(mm$Date)
summary(mm$Date)
```

Great. Looks like we're ready to get started.


## Exploratory Visualisations
The excellent [GGplot2 library](http://docs.ggplot2.org/current/) implements the [Grammar of Graphics](http://www.springer.com/statistics/computational+statistics/book/978-0-387-24544-7).

Note the pause between charts is intended for interactive use.

```{r  fig.width=7, fig.height=3, results =FALSE }
library(ggplot2)

for(var in colnames(mm)[2:ncol(mm)]) {
  print(qplot(Date, mm[,var], data=mm, geom="path", ylab=var))
  key <- readline("Press <return> for next plot")
  if(key!="") { break }
}
```

To plot these trends on the same chart we must first [normalise the data](http://en.wikipedia.org/wiki/Database_normalization). The `reshape2` library provides the `melt` function for this purpose.

```{r}
library(reshape2)
mm.melted <- melt(mm, id.vars="Date")
head(mm.melted)
```
```{r fig.width=7, fig.height=14}
ggplot(mm.melted, aes(Date, value)) + geom_path(na.rm=T) + geom_smooth(method="loess", na.rm=T) + facet_wrap(~ variable, ncol=1, scales="free_y")
```

We can create bivariate plots to investigate the relationships between variables.
```{r}
ggplot(mm, aes(Calories, Steps)) + geom_path(aes(colour=as.integer(row.names(mm))), alpha=0.5, na.rm=T)
ggplot(mm, aes(Stress, MoodInt)) + geom_path(aes(colour=as.integer(row.names(mm))), alpha=0.5, na.rm=T)
```
